### 2015-05-05 internal_review_response.Rnw

\setlength{\parskip}{5mm}
\setlength{\parindent}{0in}

\begin{document}
\raggedright

% \title{}
% \author{}
% \maketitle

{\it Response to review from Dr. Jeff Hollister, ``SWMPr: An R Package for Retrieving, Organizing, and Analyzing Environmental Data for Estuaries'', Beck, MW. May 2015.}

Response to reviewer comments are provided in italics.  The original review is included as a pdf with markup comments. The original pages and line numbers are noted below.

@@

Page 2: identifying the signal from the

{\it Changed.}

Page 2: This feels like you are questioning the use of this type of data? Also, try to transition to the next paragraph. Feels a bit abrupt.

{\it Modified as follows: `A greater focus on synthesis, exploratory-based analytical techniques, and interpretation have characterized the use of data from automated monitoring programs [6, 7].  An invaluable source of monitoring data for coastal environments in the United States is provided by the National Estuarine Research Reserve System (NERRS). This network represents 28 estuarine reserves from different biogeographic regions that were chosen to address multiple goals for long-term research, monitoring, education, and stewardship in support of coastal management.'}

@@

Page 2: both for

{\it Changed.}

Page 2: Why just abiotic? You should add a bit about this. Are the other data not available, just haven't done it yet? If the later, let me know, that'd be fun to work on!

{\it The description in the text is vague and I've modified for clarity: `Only the abiotic data are monitored continuously with automated sensor networks, whereas the remaining elements involve field surveys or mapping products that differ between reserves given site-specific requirements.  As such, the SWMPr package is developed for the continuous abiotic monitoring network that represents a majority of the SWMP data and, consequently, the most challenging to evaluate.  '}

@@

Page 3: On CRAN now, correct? IF so update installation directions.

{\it Installation instructions were updated}

Page 3: Would bet a lot of readers won't have any idea what this means. Since the swmpr object is central to the package, maybe a section describing its structure and a line or two on S3 might be a good idea.

{\it Agreed.  This material was already in the text but not as a separate section.  I added a section describing the object structure, including the following text for clarity: `The \texttt{swmpr} object class was created for use with the organizing and analyzing functions.  This object-oriented approach is standard for R (i.e., the S3 object  system, [16]), such that specific methods for generic functions are developed for the object class.}

@@

Page 4: Why only 100? web service limit?

{\it I've modified all functions that have limits on the number of records that can be downloaded.  Text was amended to remove any description of the limits, with the exception of the following: `Due to rate limitations on the CDMO server, the retrieval functions return a limited number of records with each request.  However, the SWMPr functions use the native CDMO web services iteratively (i.e., within a loop) to obtain the desired time series.  Download time can be excessive for longer time series.'}

@@

Page 4: Same as above. If there are service limitations should outline those. Just saw this explained below!

{\it Same as above.}

Page 4: This is going to be a tough one to keep relevant. What's the reason for including it?

{\it Agreed, the description of this function was removed from the manuscript but it is retained in the package and briefly noted in the supporting information.  I think it's useful to include but shouldn't be advertised as a primary means of importing data.}

@@

Page 5: Unless it is necessary, I wouldn't even mention this.

{\it Omitted.}

Page 5: Would be nice to have the decompression as part of the function.

{\it The function was modified to include this feature.  The text was updated accordingly `The \texttt{import\_local} function can be used to import files directly from the zipped folder or after the folder is decompressed.'}

@@

Page 7: Not sure how best to refer to these S3 methods. You do it both ways (with .smpr in the table) and without in the text. Could keep it that way if you add a line to table caption. I bring it up, because I think new R users would be confused by this (I was for awhile...)

{\it Agreed, I've removed the method identifier from each function in the tables to avoid confusion.}

Page 9: Does PLOS allow Boxes? Like Box 3. Example of qaqc() funciton usage... That would be a good way to show these.

{\it I've checked a few papers from PLOS One that describe software packages and most follow the format I've used, i.e., command prompt followed by code, no identifying box.}

@@

Page 9: quote

{\it Added.}

Page 10: A section on package dependencies might be a good addition.

{\it I've added a footnote that directs readers to the package index page on CRAN where the dependencies and package imports can be viewed.  I'm not sure it's worth describing each in detail since this is readily available online and in the source code.}

@@

Page 19: Yank this. Goes without saying.

{\it Removed.}

Page 20: Some of the figures have code that directly preceeds it. Other figures (like Fig 8) it is a bit more difficult to link code to figure. Would be nice to provide a link to code for all the figures. This way a reader could download a single script to reproduce the figures.

{\it I've added an R script to the supporting information that reproduces the figures/tables in the manuscript, including a link to the data.}

@@

Page 20: Are these examples necessary? Feel kinda tacked on and why have links as supplementary info and not part of the text? If you are going to keep these, I would flesh this out a bit more, add in some screen captures as figures and links directly in the text.

{\it I've added links to each app as footnotes and also included a figure for the SWMP comparison map.  The following text was added to the end of the paragraph: `Both of the apps provide previously unavailable utilities to interact with SWMP data using an intuitive online interface.  This format may be more appealing for individuals with limited experience using R, while simultaneously illustrating what is possible using functions from the package.  The applications have been used extensively with over 300 hours noted in a single month.'}

### 2015-05-05 swmpr_manu.Rnw
internal review revisions done

An important monitoring network for coastal environments in the United States is the \acl{NERRS} (\acs{NERRS})\footnote{\url{http://www.nerrs.noaa.gov/}}.  This network represents 28 estuarine reserves from different biogeographic regions that were chosen to address multiple goals for long-term research, monitoring, education, and stewardship in support of coastal management.  As part of this effort, the \acf{SWMP}\footnote{\url{http://nerrs.noaa.gov/RCDefault.aspx?ID=18}} was implemented in 1995 at over 140 stations at the reserves to provide a robust, long-term monitoring system for water quality, weather, and land-use/habitat change.  The \ac{SWMP} network has provided a continuous source of data collected at near real-time with the intent to evaluate natural and anthropogenic causes of spatiotemporal variation in environmental condition and ecosystem function.  These data have been applied both for evaluations of relevant characteristics at individual reserves (eg., \cite{Bulthius95,Dix08}) and differences between reserves (e.g., ecosystem metabolism \cite{Caffrey03,Caffrey04}, tidal characteristics \cite{Sanger02}, dissolved oxygen \cite{Wenner04}).  However, no cross-reserve comparisons have been conducted within the last decade despite the online availability of current \ac{SWMP} data.  \ac{NERRS} researchers and staff have also expressed a need for quantitative analysis tools to evaluate trends in water quality time series given the quantity and quality of data provided by \ac{SWMP} \cite{SWMP14}.

@@

The second approach for data retrieval is to use the \texttt{import\_local} function to import data into R that are locally available after downloading from \ac{CDMO}.  This approach is most appropriate for large, specific data requests. The \texttt{import\_local} function is designed for data from the zip downloads\footnote{\url{http://cdmo.baruch.sc.edu/aqs/zips.cfm}} feature in the advanced query section of the \ac{CDMO}. The zip downloads feature can be used to obtain data from multiple stations in one request.  The downloaded data will be in a compressed folder that includes multiple .csv files by year for a given data type (e.g., apacpwq2002.csv, apacpwq2003.csv, apacpnut2002.csv, etc.).  The \texttt{import\_local} function can be used to import files directly from the zipped folder or after the folder is decompressed

@@

The \texttt{swmpr} object class was created for use with the organizing and analyzing functions.  This object-oriented approach is standard for R (i.e., the S3 object  system, \cite{Wickham14}), such that specific methods for a generic function are developed for the object class.  A \texttt{swmpr} object also secondarily inherits methods from the \texttt{data.frame} class, such that common \texttt{data.frame} methods also apply to \texttt{swmpr} objects.  Available methods for the \texttt{swmpr} class are described below and can also be viewed:

@@

Detailed exploratory graphics are also useful for evaluating general trends in observed data.  Several graphics showing seasonal and annual trends for a single SWMP parameter can be obtained using the \texttt{plot\_summary} function (\Cref{fig:summary_ex}).  The plots include monthly distributions, monthly anomalies, and annual anomalies in multiple formats.  Anomalies are defined as the difference between the monthly or annual averages from the grand mean for the parameter.  An interactive web application \cite{Chang15} that uses this function is available for viewing results of any parameter at all \ac{SWMP} sites (see the \nameref{supp_info}).

@@

The second and third examples are two interactive web applications \cite{Chang15} created using the SWMPr package that illustrate summaries and comparisons of \ac{SWMP} data (see the \nameref{supp_info}).  The first web application evaluates trends in \ac{SWMP} data within and between sites using an interactive map.  Trends between reserves can be viewed using the map, whereas trends at individual sites can be viewed by clicking on a map location.  Site-level trends are shown below the map with a simple linear regression to show an increase or decrease in values over time.  Trends on the map at each station are plotted as circles that identify the direction and significance of the trend, such that larger points with darker colors indicate a strong trend.  The trend direction is blue for decreasing and red for increasing with time.  The second application provides graphical summaries of water quality, weather, or nutrient station data at individual stations using the \texttt{plot\_summary} function. The drop down menus can be used to select the station, date range, and parameter for plotting. The data used for each application are similar to those used to estimate ecosystem metabolism described above and are available in the \nameref{supp_info}.

@@

\subsection*{Trends in \ac{SWMP} parameters}
\label{swmp_trends}
Evaluate trends within and between reserves using an interactive map: \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}

\subsection*{Monthly and annual summary of \ac{SWMP} parameters}
\label{swmp_summary}
Summarize trends in \ac{SWMP} parameters for an individual site: \href{https://beckmw.shinyapps.io/swmp_summary/}{https://beckmw.shinyapps.io/swmp\_summary/}

@@

Quickly access all \ac{SWMP} data from 1995 through 2014. See the help file for \texttt{import\_remote} for more information: \href{https://s3.amazonaws.com/swmpalldata/}{https://s3.amazonaws.com/swmpalldata/}

@@

Source code as an R script for reproducing figures and tables in the manuscript: \href{}{}

>>>>

An invaluable source of monitoring data for coastal environments in the United States is provided by the \acl{NERRS} (\acs{NERRS})\footnote{\url{http://www.nerrs.noaa.gov/}}.  This network represents 28 estuarine reserves from different biogeographic regions that were chosen to address multiple goals for long-term research, monitoring, education, and stewardship in support of coastal management.  As part of this effort, the \acf{SWMP}\footnote{\url{http://nerrs.noaa.gov/RCDefault.aspx?ID=18}} was implemented in 1995 at over 140 stations across the reserves to provide a robust, long-term monitoring system for water quality, weather, and land-use/habitat change.  The \ac{SWMP} network has provided a continuous source of data collected at near real-time with the intent to evaluate natural and anthropogenic causes of spatiotemporal variation in environmental condition and ecosystem function.  These data have been applied both for evaluations of relevant characteristics at individual reserves (eg., \cite{Bulthius95,Dix08}) and differences between reserves (e.g., ecosystem metabolism \cite{Caffrey03,Caffrey04}, tidal characteristics \cite{Sanger02}, dissolved oxygen \cite{Wenner04}).  However, no cross-reserve comparisons have been conducted within the last decade despite the online availability of current \ac{SWMP} data.  \ac{NERRS} researchers and staff have also expressed a need for quantitative analysis tools to evaluate trends in water quality time series given the quantity and quality of data provided by \ac{SWMP} \cite{SWMP14}.

@@

The second approach for data retrieval is to use the \texttt{import\_local} function to import data into R that are locally available after downloading from \ac{CDMO}.  This approach is most appropriate for large, specific data requests. The \texttt{import\_local} function is designed for data from the zip downloads\footnote{\url{http://cdmo.baruch.sc.edu/aqs/zips.cfm}} feature in the advanced query section of the \ac{CDMO}. The zip downloads feature can be used to obtain data from multiple stations in one request.  The downloaded data will be in a compressed folder that includes multiple .csv files by year for a given data type (e.g., apacpwq2002.csv, apacpwq2003.csv, apacpnut2002.csv, etc.).  The \texttt{import\_local} function can be used to import files directly from the zipped folder or after the folder is decompressed.

@@

The \texttt{swmpr} object class was created for use with the organizing and analyzing functions.  This object-oriented approach is standard for R (i.e., the S3 object  system, \cite{Wickham14}), such that specific methods for generic functions are developed for the object class.  A \texttt{swmpr} object also secondarily inherits methods from the \texttt{data.frame} class, such that common \texttt{data.frame} methods also apply to \texttt{swmpr} objects.  Available methods for the \texttt{swmpr} class are described below and can also be viewed:

@@

Detailed exploratory graphics are also useful for evaluating general trends in observed data.  Several graphics showing seasonal and annual trends for a single SWMP parameter can be obtained using the \texttt{plot\_summary} function (\Cref{fig:summary_ex}).  The plots include monthly distributions, monthly anomalies, and annual anomalies in multiple formats.  Anomalies are defined as the difference between the monthly or annual averages from the grand mean for the parameter.  An interactive web application \cite{Chang15} that uses this function is available for viewing results of any parameter at all \ac{SWMP} sites (see the \nameref{swmp_apps} section).

@@

The second and third examples are two interactive web applications \cite{Chang15} created using the SWMPr package that illustrate summaries and comparisons of \ac{SWMP} data.  The first web application\footnote{Evaluate trends within and between reserves using an interactive map: \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}} evaluates trends in \ac{SWMP} data within and between sites using an interactive map (\cref{fig:swmp_comp}).  Trends between reserves can be viewed using the map, whereas trends at individual sites can be viewed by clicking on a map location.  Site-level trends are shown below the map with a simple linear regression to show an increase or decrease in values over time.  Trends on the map at each station are plotted as circles that identify the direction and significance of the trend, such that larger points with darker colors indicate a strong trend.  The trend direction is blue for decreasing and red for increasing with time.  The second application\footnote{Summarize trends in \ac{SWMP} parameters for an individual site: \href{https://beckmw.shinyapps.io/swmp_summary/}{https://beckmw.shinyapps.io/swmp\_summary/}} provides graphical summaries of water quality, weather, or nutrient station data at individual stations using the \texttt{plot\_summary} function (output is identical to \cref{fig:summary_ex}). Drop down menus can be used to select the station, date range, and parameter for plotting. The data used for each application are similar to those used to estimate ecosystem metabolism described above and are available in the \nameref{supp_info}.  Both of the apps provide previously unavailable utilities to interact with \ac{SWMP} data using an intuitive online interface.  This format may be more appealing for individuals with limited experience using R, while simultaneously illustrating what is possible using functions from the package. The applications have been used extensively with over 300 hours noted in a single month.

\begin{figure}
\begin{center}
\includegraphics[width = \textwidth]{figure/swmp_comp.pdf}
\caption{Online application for comparing trends in \ac{SWMP} data parameters using an interactive map.  Link: \href{https://beckmw.shinyapps.io/swmp_comp}{https://beckmw.shinyapps.io/swmp\_comp}}
\label{fig:swmp_comp}
\end{center}
\end{figure}

@@

@@

All \ac{SWMP} data from 1995 through 2014, one file per site. See the help file for \texttt{import\_remote} for more information: \href{https://s3.amazonaws.com/swmpalldata/}{https://s3.amazonaws.com/swmpalldata/}

@@

Source code as an R script for reproducing figures and tables in the manuscript: \href{https://github.com/fawda123/swmpr_manu/blob/master/swmpr_manu_code.R}{https://github.com/fawda123/swmpr\_manu/blob/master/swmpr\_manu\_code.R}

\subsection*{Data to reproduce figures/tables}

Data used to create the figures and tables.  Files are in the .RData format: \href{https://github.com/fawda123/swmpr_manu/tree/master/data}{https://github.com/fawda123/swmpr\_manu/tree/master/data}
